Mixed effect tweedie model in TMB

### 2019-05-17 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
misc changes to tmb-sensor-explore

Load data
```{r}
all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST-all.rds"))
# all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST
# glimpse(all_sensor)
# 
# replace obviously faulty depth data with NA
all_sensor$depth_m[all_sensor$depth_m < 10] <- NA

all_depth <- all_sensor %>%
  # filter missing location data and trial year
  dplyr::filter(!is.na(latitude), !is.na(longitude)) %>%

@@

d_trawl <- all_depth$data %>% 
  # filter(ssid != 1) %>%
  # filter(ssid != 3) %>%
  # filter(ssid != 4) %>%
  # filter(ssid != 16) %>%
  gfplot:::scale_survey_predictors()

@@

# remove obvious sensor fail, d_trawl %>% filter(ssid==4, temperature_c >12), fishing event id = 481861
d_trawl <- d_trawl[d_trawl$fishing_event_id != 481861, ]
d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
d_trawl <- d_trawl[!is.na(d_trawl$meanSST), ]
d_trawl <- mutate(d_trawl,
  DOY = lubridate::yday(date),
  DOY_scaled = arm::rescale(lubridate::yday(date)),
  #depth_scaled = arm::rescale(log(depth_m)),
  SST_scaled = arm::rescale(meanSST),
  #depth_scaled2 = depth_scaled^2,
  depth_scaled3 = depth_scaled^3,
  shallow = ifelse(depth>35,0,1)
)
# d_trawl[is.na(d_trawl$shallow), ]
# d_trawl[is.na(d_trawl$SST_scaled), ]

@@

```{r}
ggplot(dplyr::filter(d_trawl, depth <35), aes(meanSST,temperature_c, colour = -depth)) + geom_point()
```

@@

@@

```{r}
spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 650)
plot_spde(spde)
```

```{r}
m_temp3 <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year) + shallow*poly(DOY_scaled,2), 
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

m_temp3b <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year), 
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

m_temp3
m_temp3b
```

@@

predictions3 <- predict(m_temp3)
predictions3b <- predict(m_temp3b)
#predictions3b <- predict(m_temp3_nosst)

predictions3$data$residuals <- residuals(m_temp3)
predictions3b$data$residuals <- residuals(m_temp3b)

nd <- data.frame(sst_resid = predictions3$data$residuals, nosst_resid = predictions3b$data$residuals, shallow = d_trawl$shallow, depth = d_trawl$depth)

ggplot(filter(nd, shallow == 1), aes(nosst_resid,sst_resid, colour = as.factor(shallow))) + 
  geom_point() + geom_abline(intercept = 0, slope = 1)
```
Residual plots:
```{r}
ggplot(predictions3$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  ylim(-8,8) +
  xlim(2,15) +
  facet_wrap(~year)
ggplot(predictions3b$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  ylim(-8,8) +
   xlim(2,15) +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  ylim(-7,8) +
  facet_wrap(~year)
ggplot(predictions3b$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  ylim(-7,8) +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3$data, aes(depth_scaled, residuals)) +
  geom_point() +
  ylim(-8,8) +
  geom_smooth()
ggplot(predictions3b$data, aes(depth_scaled, residuals)) +
  geom_point() +
  ylim(-8,8) +
  geom_smooth()
```

>>>>

Load data
```{r}
all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST-all.rds"))
# all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST
# glimpse(all_sensor)
# 
# replace obviously faulty depth data with NA
#all_sensor$depth_m[all_sensor$depth_m < 10] <- NA # 5,139

all_depth <- all_sensor %>%
  #filter(depth_max>10) %>% # maybe we should filter incase missing depth values also indicate a sensor fail?
  # filter missing location data and trial year
  dplyr::filter(!is.na(latitude), !is.na(longitude)) %>%

@@

d_trawl <- all_depth$data %>% 
  # remove obvious sensor fails, d_trawl %>% filter(depth_max<10), fishing event ids = 481861, 2179089
  filter(depth_max>10) %>% 
  # filter(ssid != 1) %>%
  # filter(ssid != 3) %>%
  # filter(ssid != 4) %>%
  # filter(ssid != 16) %>%
  gfplot:::scale_survey_predictors()

@@

d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
d_trawl <- d_trawl[!is.na(d_trawl$meanSST), ]
d_trawl <- mutate(d_trawl,
  DOY = lubridate::yday(date),
  DOY_scaled = arm::rescale(lubridate::yday(date)),
  DOY_scaled2 = arm::rescale(lubridate::yday(date))^2,
  #depth_scaled = arm::rescale(log(depth_m)),
  SST_scaled = arm::rescale(meanSST),
  #depth_scaled2 = depth_scaled^2,
  depth_scaled3 = depth_scaled^3,
  shallow = ifelse(depth>35,0,1),
    deep = ifelse(depth<35,0,1)
)
# d_trawl[is.na(d_trawl$shallow), ]
# d_trawl[is.na(d_trawl$SST_scaled), ]

@@

```{r}
ggplot(dplyr::filter(d_trawl1, shallow==1), aes(SST_scaled,temperature_c, colour = -depth)) + geom_point()
```

@@

```{r}
ggplot(d_trawl, aes(log(depth),temperature_c, colour = depth_max)) + geom_point() + facet_wrap(~ssid)
```

@@

```{r}
d_trawl1 <- d_trawl %>% filter(ssid == 3)
spde <- make_spde(d_trawl1$X, d_trawl1$Y, n_knots = 150)
plot_spde(spde)
sum(d_trawl1$shallow)
```

```{r}
d_trawl1$year_scaled <- arm::rescale(d_trawl1$year)
d_trawl1$x_scaled <- arm::rescale(d_trawl1$X)
d_trawl1$y_scaled <- arm::rescale(d_trawl1$Y)

m_temp_SST <- sdmTMB(d_trawl1,
  temperature_c ~ year_scaled + poly(SST_scaled,3)*depth_scaled , #+ SST_scaled*x_scaled + SST_scaled*y_scaled
  #time_varying = ~ 0 + depth_scaled2,
  ar1_fields = FALSE, include_spatial = FALSE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = FALSE
)

p_SST <- predict(m_temp_SST)

p_SST$residuals <- residuals(m_temp_SST)

ggplot(p_SST, aes(depth_scaled, residuals)) +
  geom_point(alpha=0.4) +
  ylim(-8,8) +
  geom_smooth()

```{r}
library(gbm)

m <- gbm( temperature_c ~ depth_scaled + X + Y, data = d_trawl1, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)

m2 <- gbm( temperature_c ~ SST_scaled + depth_scaled + X + Y, data = d_trawl1, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)

plot(m,i.var=2)
plot(m,i.var=3)
plot(m,i.var=4)
plot(m,i.var=5)

plot(m2,i.var=1)
plot(m2,i.var=2)
plot(m,i.var=3)
plot(m,i.var=4)

plot(m2,i.var=1:2)
plot(m2,i.var=c(1,3))
plot(m2,i.var=c(1,4))
d_trawl$r <- predict(m, n.trees = 2000) - d_trawl1$temperature_c

ggplot(d_trawl, aes(depth_scaled, r)) +
  geom_point(alpha=0.4) +
  ylim(-8,8) +
  geom_smooth()
```

@@

p_SST <- predict(m_temp_SST)
p <- predict(m_temp)
#predictions3b <- predict(m_temp3_nosst)

p_SST$residuals <- residuals(m_temp_SST)
p$residuals <- residuals(m_temp)

# nd <- data.frame(sst_resid = predictions3$data$residuals, nosst_resid = predictions3b$data$residuals, shallow = d_trawl$shallow, depth = d_trawl$depth)
# 
# ggplot(filter(nd, shallow == 1), aes(nosst_resid,sst_resid, colour = as.factor(shallow))) + 
#   geom_point() + geom_abline(intercept = 0, slope = 1)
```
Residual plots:
```{r}
ggplot(p_SST, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  ylim(-8,8) +
  xlim(2,15) +
  facet_wrap(~year)
ggplot(p, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  ylim(-8,8) +
   xlim(2,15) +
  facet_wrap(~year)
```

```{r}
ggplot(p_SST, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  ylim(-7,8) +
  facet_wrap(~year)
ggplot(p, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  ylim(-7,8) +
  facet_wrap(~year)
```

```{r}
p_SST$depth_mean <- exp(p_SST$depth_scaled*0.66+5)
glimpse(d_trawl)
ggplot(d_trawl, aes(depth_scaled, r)) +
  geom_point(alpha=0.4) +
  ylim(-8,8) +
  geom_smooth()
```

```{r}
p1 <- ggplot(p_SST, aes(depth_scaled, residuals)) +
  geom_point() +
  ylim(-8,8) +
  geom_smooth()
p2 <- ggplot(p, aes(depth_scaled, residuals)) +
  geom_point() +
  ylim(-8,8) +
  geom_smooth()
gridExtra::grid.arrange(p1, p2, nrow = 2)
```
